R for Clinical Study Reports and Submission

### 2021-11-08 index.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
options(dplyr.summarise.inform = FALSE)
file.copy("static/CNAME", "_book", overwrite = TRUE)
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "#>", echo = TRUE, message = FALSE)
options(dplyr.summarise.inform = FALSE)
file.copy("static/CNAME", "_book", overwrite = TRUE)
fs::dir_copy("slides/", "_book/slides", overwrite = TRUE)
```

### 2021-11-08 intro.Rmd
Minor typo fix

In clinical trial development, source code needs to be developed and maintained to generate and deliver 
Study Data Tabulation Model (SDTM), Analysis Dataset Model (ADaM) datasets and tables, listings, and figures (TLFs).
A typical example is a Phase 3 clinical trial where hundreds of TLFs are required for submission. 
Considering the amount of programs required for such an effort, 
a consistent and well-defined folder structure is crucial in 
managing a clinical trial analysis and reporting (A&R) project.

We recommend using 
[R package folder structure](https://github.com/rstudio/cheatsheets/raw/master/package-development.pdf) 
to organize all the A&R related source code and documentation for a clinical trial A&R project. 
The R package folder structure is well defined and widely used in R community through repositories (e.g. CRAN).

The consistent approach of using R package folder structure 
simplifies the communication for all developers within and across organizations.

- For a new R developer, 
it is an essential step to develop R packages 
when you want to share your work with others. 
You will learn one folder structure that is widely used 
in the R community with outstanding tutorials and tools for free.

- For an experienced R developer, there is a minimal learning curve.

- For an organization, it simplifies the process, tool, template, 
and training development, because a unified folder structure is used 
to develop and maintain standard tool and analysis projects.

The workflow around an R package can also improve 
the tractability and reproducibility for an analysis project [@marwick2018packaging].

In addition, R package folder structure is also recommended 
to develop Shiny apps as discussed 
in [Chapter 20 of Master Shiny](https://mastering-shiny.org/scaling-packaging.html) book 
and [Engineering Production-Grade Shiny Apps](https://engineering-shiny.org/golem.html) book.

@@

- Part 2, "Clinical Trial Project" provides general information 
with examples to manage a clinical trial A&R project. 
In this part, we assume a reader is a project lead with experience in R package development. 
Some good reference includes [R packages](https://r-pkgs.org/) and 
[The tidyverse style guide](https://style.tidyverse.org/).

- Part 3, "eCTD Submission package"
provides general information in preparing 
submission package related to clinical study report (CSR) in 
electronic Common Technical Document (eCTD). 
In this part, we assume a reader is a project lead of a clinical project 
with experience in R package development and submission.

@@

We share the same philosophy described in 
[Section 1.1 of R Packages book](https://r-pkgs.org/intro.html?q=Marwick#intro-phil) 
and quote here.

>>>>

In clinical trial development, source code needs to be developed and maintained to generate and deliver
Study Data Tabulation Model (SDTM), Analysis Dataset Model (ADaM) datasets and tables, listings, and figures (TLFs).
A typical example is a Phase 3 clinical trial where hundreds of TLFs are required for submission.
Considering the number of programs required for such an effort,
a consistent and well-defined folder structure is crucial in
managing a clinical trial analysis and reporting (A&R) project.

We recommend using the
[R package folder structure](https://github.com/rstudio/cheatsheets/raw/master/package-development.pdf)
to organize all the A&R-related source code and documentation for a clinical trial A&R project.
The R package folder structure is well defined and widely used in the R community through repositories (e.g., CRAN).

The consistent approach of using the R package folder structure
simplifies the communication for all developers within and across organizations.

- For a new R developer, it is an essential step to develop R packages
  when you want to share your work with others.
  You will learn one folder structure widely used in the R community
  with outstanding tutorials and tools for free.
- For an experienced R developer, there is a minimal learning curve.
- For an organization, it simplifies the process, tool, template,
  and training development, because a unified folder structure is used
  to develop and maintain standard tool and analysis projects.

The workflow around an R package can also improve
the traceability and reproducibility of an analysis project [@marwick2018packaging].

We will revisit folder structure when we discuss project management for a clinical trial project.

In addition, R package folder structure is also recommended
to develop Shiny apps as discussed
in [Chapter 20 of the Master Shiny book](https://mastering-shiny.org/scaling-packaging.html)
and the [Engineering Production-Grade Shiny Apps](https://engineering-shiny.org/golem.html) book.

@@

- Part 2, "Clinical trial project" provides general information
  with examples to manage a clinical trial A&R project.
  In this part, we assume a reader is a project lead with experience in R package development.
  Some good references include [R Packages](https://r-pkgs.org/) and
  [The tidyverse style guide](https://style.tidyverse.org/).

- Part 3, "eCTD submission package" provides general information in preparing
  submission packages related to clinical study report (CSR) in
  electronic Common Technical Document (eCTD).
  In this part, we assume a reader is a project lead of a clinical project
  with experience in R package development and submission.

@@

We share the same philosophy described in
[Section 1.1 of the R Packages book](https://r-pkgs.org/intro.html#intro-phil)
and quote here.

### 2021-11-08 slides/workshop-slides.Rmd
fix typo
rename `adam_data/` as `data-adam/`
set comment = "#>"

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

@@

- Source data: <https://github.com/elong0527/r4csr/tree/master/adam_data>

@@

`r2rtf` only focuses on **table format**. 
Data manipulation and analysis shall be handled by other R packages. (e.g. `tidyverse`)

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "#>", echo = FALSE)
```

@@

- Source data: <https://github.com/elong0527/r4csr/tree/master/data-adam>

@@

`r2rtf` only focus on **table format**.
Data manipulation and analysis shall be handled by other R packages. (e.g. `tidyverse`)

### 2021-11-08 submission-overview.Rmd
address Nan's comments
use sentence case instead of title case for headings; run styler on Rmds

# (PART) eCTD Submission {-}

# Overview

@@

> "Sponsors should provide the software programs used to create all ADaM
datasets and **generate tables and figures associated with primary and
secondary efficacy analyses**. Furthermore, sponsors should submit software
programs used to generate additional information included in Section 14
CLINICAL STUDIES of the Prescribing Information (PI)26 if applicable.
**The specific software utilized should be specified in the ADaM Reviewer's Guide (ADRG)**.
The main purpose of requesting the submission of these programs
**is to understand the process by which the variables for the respective
analyses were created and to confirm the analysis algorithms**.
Sponsors should submit software programs in **ASCII text format**;
however, executable file extensions should not be used."

In Chapter \@ref(submission-package), we will focus on the approach
to prepare proprietary R packages and analysis code into proper
formats for submission.

In Chapter \@ref(running-environment), we will discuss
the recommendations to make the R code running environment
reproducible for dry run tests and reviews.

>>>>

# (PART) eCTD submission {-}

# Overview {#submission-overview}

@@

> "Sponsors should provide the software programs used to create all ADaM
datasets and **generate tables and figures associated with primary and
secondary efficacy analyses**. Furthermore, sponsors should submit software
programs used to generate additional information included in Section 14
CLINICAL STUDIES of the Prescribing Information (PI)26 if applicable.
**The specific software utilized should be specified in the ADRG**.
The main purpose of requesting the submission of these programs
**is to understand the process by which the variables for the respective
analyses were created and to confirm the analysis algorithms**.
Sponsors should submit software programs in **ASCII text format**;
however, executable file extensions should not be used."

Chapter \@ref(submission-package) will focus on preparing
proprietary R packages and analysis code into proper
formats for submission.

Chapter \@ref(running-environment) will discuss
the recommendations to make the R code running environment
reproducible for dry run tests and reviews.

### 2021-11-08 tlf-overview.Rmd

For simplicity, we have downloaded all these datasets into the `adam_data` folder of
this project and converted them from the `.xpt` format to the `.sas7bdat` format.

>>>>

For simplicity, we have downloaded all these datasets into the `data-adam/`
folder of this project and converted them from the `.xpt` format to
the `.sas7bdat` format.
